---
title: "Disturbance"
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output:
  html_document: 
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document: default
  word_document: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()) ; invisible(gc())
library(knitr)
library(parallel)
library(TROLL)
library(RconTroll)
library(ggplot2)
library(cowplot)
library(entropart)
library(plotly)
cores <- 16
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
    cache = T, cache.lazy = F)
```

Built with `RconTroll` package version `r packageVersion('RconTroll')` available at https://gitlab.com/sylvain.schmitt/RconTroll and `TROLL` package version `r packageVersion('TROLL')` available at https://gitlab.com/sylvain.schmitt/TROLL with `r R.Version()$version.string` environment.

# Design of experiment

## Space of experiments

```{r Sample, include=FALSE}
# Init
path <- '~/Documents/ECOFOG/Results/disturbance/'
set.seed(42)
n <- 20
rep <- 10^3
dist <- 0:3*0.25
rich <- 5^(1:3)
# # Coms
# species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt',
#                       header=TRUE, dec=".", sep="",
#                       row.names = 1)[c('LMA', 'wsg', 'hmax', 'dmax')]
# proba <- ks::kde(scale(species), eval.points = scale(species))$estimate
# proba <- 1/proba
# coms <- lapply(as.list(rich), function(r)
#   replicate(n*rep, species[sample(row.names(species), r, F, proba),],
#             simplify = F))
# names(coms) <- rich ; rm(proba)
# # Dists
# CWM <- lapply(coms, function(x) do.call('rbind',
#                                         lapply(x, function(y) apply(y,2, mean))))
# Centroids <- lapply(CWM, function(x) apply(x, 2, mean))
# Dist <- data.frame(mapply(function(x,y)
#   apply(x, 1, function(z) sqrt(sum((z - y)^2))), x = CWM, y = Centroids))
# names(Dist) <- paste0('dist.', rich) ; rm(CWM, Centroids)
# # Convex hull
# CH <- lapply(coms, function(x)
#   do.call('rbind', lapply(x, function(y)
#     try(geometry::convhulln(y, options = 'FA')$vol))))
# CH$`5` <- as.numeric(CH$`5`)
# CH <- data.frame(do.call('cbind', CH))
# names(CH) <- paste0('CH.', rich)
# # Sample
# Sample <- cbind(Dist, CH)
# Sample <- list(
#   `5` = Sample[c(1,4)],
#   `25` = Sample[c(2,5)],
#   `125` = Sample[c(3,6)]
# )
# Sample <- lapply(Sample, function(x) x[x[,1] < sd(x[,1]),])
# Sample <- lapply(Sample, function(x) {x$num <- row.names(x) ; return(x)})
# Scale <- lapply(Sample, function(x)
#   seq(from = min(x[,2], na.rm = T), max(x[,2], na.rm = T), length.out = n))
# Sample <- mapply(function(x,y) do.call('rbind', sapply(y, function(z)
#   x[which.min(abs(x[,2] - z)) ,], simplify = F)),  x = Sample, y = Scale, SIMPLIFY = F)
# Sel <- lapply(Sample, function(x) x$num)
# Sel <- mapply(function(x,y) x[as.numeric(y)], x = coms, y = Sel, SIMPLIFY = F)
# Sel <- lapply(Sel, function(x) lapply(x, row.names))
# names(Sel) <- rich
sim <- as.vector(sapply(rich, function(r) paste(r, 1:n, sep = '.')))
# Sel <- unlist(Sel, recursive = F)
# names(Sel) <- sim ; rm(Scale, CH, Dist, coms)
# save(Sel, Sample, file = file.path(path, 'SpeciesSel2.Rdata'))
load(file.path(path, 'SpeciesSel2.Rdata'))
```

The space of experiments encompass following axes (with their abbreviations and values) :

- **Disturbance:** $dist \in \left\{ `r paste(dist)` \right\}$
- **Richness:** $rich \in \left\{ `r paste(rich)` \right\}$
- **Community convexhull volume:** $ch$

Community hypervolume will depend on richness (higher richness implying less choices). So we can already measure possible community hypervolumes depending on the level of richness for a given community mean. To do that we sampled $`r n*rep`$ communities for each richness level. Sampling was done by attributing to each species a probability to be pick inverse of its densities in the 4-dimensional functional trait space computed with R package `ks`. Once random communities sampled, we measured each community mean and calculated community euclidean distance to the metacommunity mean. We wanted to control ecosystem functional composition with community mean. In order to do that, we decided to keep only simulated community for which community euclidean distance to metacommunity mean was smaller than the standard deviation of euclidean distances of all simulated communities. Once potential communities designated for each level of richness, we defined for each leavel of richness $`r n`$ equidistant convexhull volumes and we kept the closest community of each hypervolume level.

We obtained the followin designed of experiment (DOE) without disturbance (to replicate for each 4 levels of disturbance):

```{r DOE, fig.height=4}
DOE <- data.frame(
  Sim = sim,
  Richness = rep(rich, each = n),
  Convexhull = unlist(lapply(Sample, function(x) x[,2])),
  Distance = unlist(lapply(Sample, function(x) x[,1])))
rm(Sample, n, rep)
ggplot(DOE, aes(x = Convexhull, y = Distance, label = Sim)) +
  geom_line() + geom_point() + geom_label() +
  facet_wrap(~ Richness, scales = "free") +
  ylab('Euclidean distance to\nmetacommunity centroïd') ; rm(DOE)
```

## Pre-disturbance simulations

```{r pd Sim, include=FALSE}
# # Init
# seedrain <- 50
# dir.create(file.path(path, 'maturity'))
# invisible(sapply(sim, function(s) dir.create(file.path(path, 'maturity', s))))
# species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt',
#                       header=TRUE, dec=".", sep="")
# Sel <- lapply(Sel, function(com){
#   com <- species[which(species$Name %in% com),]
#   com$Freg <- 1/dim(com)[1]
#   return(com)
# })
# invisible(sapply(sim, function(s){
#   init(path = file.path(path, 'maturity'),
#        input = paste0(s, '.txt'),
#        species = Sel[[s]],
#        numesp = dim(Sel[[s]])[1],
#        nbiter = 12*600,
#        seedrain = seedrain)
# }))
# # Bash
# if('maturity.sh' %in% list.files(file.path(path, 'maturity')))
#    unlink(file.path(path, 'maturity', 'maturity.sh'))
# bash <- file(file.path(path, 'maturity', 'maturity.sh'), open = 'a')
# cat(' #!/bin/bash  \n\n', file = bash, append = TRUE)
# cat(' echo "maturity forest parrallel modelling in TROLL" \n\n', file = bash, append = TRUE)
# for(i in seq_len(length(sim))){
#   command <- paste0('./troll',
#                     ' -i"', sim[i], '.txt"',
#                     ' -o./', sim[i], '/', sim[i], ' &\n')
#   cat(command, file = bash, append = TRUE)
#   if(i %/% (cores/2) == i/(cores/2))
#   cat('wait\n', file = bash, append = TRUE)
# }
# cat('wait\n', file = bash, append = TRUE)
# cat(' echo "TROLL all jobs done"  \n\n', file = bash, append = TRUE)
# close(bash)
rm(bash, Sel, seedrain, i, command, species)
```

```{r pd Results, fig.height=12, fig.width=12}
# Load stack
# mature <- loadStack('~/Documents/ECOFOG/Results/disturbance/maturity/models/')
# save(mature, file = '~/Documents/ECOFOG/Results/disturbance/maturity/mature.Rdata')
# load('~/Documents/ECOFOG/Results/disturbance/maturity/mature.Rdata')

# Data prep & check
# datal <- extractData(mature)
# save(datal, file = '~/Documents/ECOFOG/Results/disturbance/maturity/matureData.Rdata')
load('~/Documents/ECOFOG/Results/disturbance/maturity/matureData.Rdata')
data <- datal$time
data$richness <- as.factor(floor(as.numeric(as.character(data$sim))))
data <- data[data$time %in% round(data$time),]
var <- sapply(c('agb', 'ba', 'n10', 'n30'), function(var)
  ggplot(data, aes_string(x = 'time', y = var,
                          label = 'sim', colour = 'richness')) +
    geom_point(),
  simplify = F)
plot_grid(plotlist = var, labels = c('AGB', 'BA', 'N10', 'N30'))

# Functional diversity
data <- datal$time
data <- data[data$time > 450,]
data <- data.frame(apply(data[c('agb','ba','n10','n30')],
                         2, function(x) tapply(x, data$sim, mean)))
data <- cbind(data.frame(sim = row.names(data)), data)
data$richness <- as.factor(floor(as.numeric(as.character(data$sim))))
species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt',
                      header=T, dec=".", sep="",
                      row.names = 1)[c('LMA', 'wsg', 'dmax', 'hmax')]
species <- species[row.names(datal$species),]
fd <- FD::dbFD(species, t(datal$species))
data <- cbind(data, do.call('cbind',
                            fd[c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ')]))
data <- cbind(data, fd$CWM)
rm(fd, species, datal)

# Correlations
corrplot::corrplot.mixed(cor(data[-c(1,6)]), order = 'hclust')

# Graphs
sapply(c('FRic', 'FDis'), function(trait){
  var <- sapply(c('agb', 'ba', 'n10', 'n30'), function(var)
  ggplot(data, aes_string(x = trait, y = var, colour = 'richness'))
  + geom_point(),
  simplify = F)
  plot_grid(plotlist = var, labels = c('AGB', 'BA', 'N10', 'N30'))
}, simplify = F)

sapply(c('LMA', 'wsg', 'dmax', 'hmax'), function(trait){
  var <- sapply(c('agb', 'ba', 'n10', 'n30'), function(var)
  ggplot(data, aes_string(x = trait, y = var, colour = 'richness'))
  + geom_point(),
  simplify = F)
  plot_grid(plotlist = var, labels = c('AGB', 'BA', 'N10', 'N30'))
}, simplify = F)

rm(data, mature, var)
```

## Monoculture simulations

```{r mm Sim}
# # Init
# seedrain <- 50 # Seedtrain constant to fix
# dir.create(file.path(path, 'monoculture'))
# dir.create(file.path(path, 'monoculture/mature'))
# sim <- sort(unique(unlist(Sel)))
# invisible(sapply(sim, function(s) dir.create(file.path(path, 'monoculture/mature', s))))
# species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt', header=TRUE, dec=".", sep="")
# Sel <- lapply(sim, function(sp){
#   com <- species[which(species$Name == sp),]
#   com$Freg <- 1
#   return(com)
# })
# names(Sel) <- sim
# invisible(sapply(sim, function(s){
#   init(path = file.path(path, 'monoculture/mature'),
#        input = paste0(s, '.txt'),
#        species = Sel[[s]],
#        numesp = 1,
#        nbiter = 12*600,
#        seedrain = seedrain)
# }))
# # Bash
# if('mature.sh' %in% list.files(file.path(path, 'monoculture/mature')))
#    unlink(file.path(path, 'monoculture/mature', 'mature.sh'))
# bash <- file(file.path(path, 'monoculture/mature', 'mature.sh'), open = 'a')
# cat(' #!/bin/bash  \n\n', file = bash, append = TRUE)
# cat(' echo "mature monoculture forest parrallel modelling in TROLL" \n\n',
#     file = bash, append = TRUE)
# for(i in seq_len(length(sim))){
#   command <- paste0('./troll',
#                     ' -i"', sim[i], '.txt"',
#                     ' -o./', sim[i], '/', sim[i], ' &\n')
#   cat(command, file = bash, append = TRUE)
#   cat(paste0(' echo "', sim[i], ' Done" \n'), file = bash, append = TRUE)
#   if(i %/% (cores/2) == i/(cores/2))
#   cat('wait\n', file = bash, append = TRUE)
# }
# cat('wait\n', file = bash, append = TRUE)
# cat(' echo "TROLL all jobs done"  \n\n', file = bash, append = TRUE)
# close(bash)
```

```{r mm Results, fig.height=12, fig.width=12}
# Load stack 
# mm <- loadStack('~/Documents/ECOFOG/Results/disturbance/monoculture/mature/')
# save(mm, file = '~/Documents/ECOFOG/Results/disturbance/monoculture/mature/mm.Rdata')

# Data prep & check
# datal <- extractData(mm)
# save(datal, file = '~/Documents/ECOFOG/Results/disturbance/monoculture/mature/mmData.Rdata')
load('~/Documents/ECOFOG/Results/disturbance/monoculture/mature/mmData.Rdata')
data <- datal$time
data <- data[data$time %in% round(data$time),]
ggplot(data, aes(x = time, y = agb, colour = sim)) +
    geom_point() + guides(colour = F)

# Functional diversity
data <- datal$time
data <- data[data$time > 450,]
data <- data.frame(apply(data[c('agb','ba','n10','n30')],
                         2, function(x) tapply(x, data$sim, mean)))
data <- cbind(data.frame(sim = row.names(data)), data)
species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt',
                      header=T, dec=".", sep="",
                      row.names = 1)[c('LMA', 'wsg', 'dmax', 'hmax')]
species <- species[row.names(datal$species),]
data$traits <- apply(species, 1, paste, collapse = ' ')
data <- cbind(data, species) # ; rm(species, datal)

# Plotly
g1 <- ggplot(data, aes(x = ba, y = agb, colour = sim, label = traits)) +
  geom_point() + guides(colour = F)
g2 <- ggplot(data, aes(x = n10, y = n30, colour = sim, label = traits)) +
  geom_point() + guides(colour = F)
g1
g2

rm(mm, data, g1, g2, p1, p2)
```

## Still an issue ?

```{r Issue}
# load('~/Documents/ECOFOG/Results/disturbance/monoculture/mature/mm.Rdata')
# m <- mm@layers$Tachigali_bracteolata
# save(m, file = '~/Documents/ECOFOG/Results/disturbance/monoculture/mature/Tachigali.Rdata')
load('~/Documents/ECOFOG/Results/disturbance/monoculture/mature/Tachigali.Rdata')
summaryPlot(m)
plot(m, what = 'age', ggplot = T)
plot(m, what = 'dbh', ggplot = T)
rm(mm)
```


## Disturbance simulations

# References