---
title: "Leaf lifespan allometry"
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
  word_document: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()) ; invisible(gc())
library(knitr)
library(parallel)
library(TROLL)
library(RconTroll)
library(reshape2)
library(ggplot2)
library(cowplot)
library(entropart)
library(plotly)
library(rstan)
library(bayesplot)
library(rstanarm)
library(readr)
library(dplyr)
library(randomForest)
library(GGally)
cores <- detectCores()
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

# Introduction

TROLL model currently compute leaf lifespan with Reich's allometry [@Reich1991a]. But we have shown that Reich's allometry is underestimating leaf lifespan for low LMA species. Moreover simulations estimated unrealistically low aboveground biomass for low LMA species. We assumed Reich's allometry underestimation of leaf lifespan for low LMA species being the source of unrealistically low aboveground biomass inside TROLL simulations. We decided to find a better allometry with @wright_worldwide_2004 GLOPNET dataset.

```{r LL alo}
rm(list = ls())
LES <- read.csv("~/Documents/ECOFOG/TROLL/inst/extdata/GLOPNET.csv", skip = 10)
LES <- LES[which(LES$BIOME == 'TROP_RF'),]
LES <- LES[which(LES$GF == 'T'),]
LES <- LES[-which(is.na(LES$log.LL)),]
data <- cbind(LES[c('Dataset', 'Species')],
              10^LES[paste0('log.',c('LL', 'LMA', 'Nmass', 'Narea',
                                     'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea'))])
names(data) <- c('Dataset', 'Species', 'LL', 'LMA', 'Nmass', 'Narea',
              'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea')
data <- droplevels(data[c('Dataset', 'Species', 'LL', 'LMA', 'Nmass')])
g <- ggplot(droplevels(data), aes(x = LMA, y = LL, size = Nmass, color = Dataset, label = Species)) +
  geom_point(alpha = 0.5) + xlab('Leaf mass per area (LMA in g/m2)') +
  ylab('Leaf lifespan (LL in months)') + 
  scale_size_continuous(name = 'Leaf nitrogen content\n(Nmass, in mg/g)')
g ; rm(LES, g)
```

Figure 1: Leaf mass per area (LMA), leaf nitrogen content (Nmass) and leaflifespan (LL). *Leaf mass per area (LMA in $g.m^{-2}$), leaf nitrogen content (Nmass, in $mg.g^-1$) and leaf lifespan (LL in $months$) are taken in GLOPNET dataset from @wright_worldwide_2004.*

# Model

We tested different models and we kept subsequant model with the best tradeoff between model complexity, convergence, likelihoof, and prediction quality (see [leaf lifespan model](LL2.html) ):
$$ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j},\,\sigma)\,$$

$$s=1,...,S_{=4}~, ~~j=1,...,n_s$$
$${\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}$$

Leaf lifespan ($LL$) is following a lognormal law with location $log({\mu_s}_j)$ and scale $\sigma$. $s$ represents the dataset used to do the fit between 1 and $S=4$, it encompass environmental and protocol variations. $j$ represents the observation for a given site $s$. Finally $X^{(1)}$ represents the LMA and $X^{(2)}$ the leaf nitrogen content (Nmass). Location ${\mu_s}_j$ is the log of the exponential from the difference between LMA and Nmass with parameters ${\beta_0}_s$, ${\beta_1}_s$, and ${\beta_2}_s$. Each ${\beta_i}_s$ is following a normal law located on $\beta_i$ with scale $\sigma_i$. Finally all $beta_i$, $\sigma$, and $\sigma_i$ are assumed without presemption following gamma laws of parameters $10^-2$, $10^-2$.

# TRY improvement

We gathered data from TRY database to evaluate and improve previous model. We requested functional traits available in Bridge dataset for species with leaf lifespan present in GLOPNET dataset (see table 1).

Table 1: Functional traits gathered with TRY.

```{r traits}
kable(traits <- data.frame(
  name = c('Thick', 'Tough', 'Nm', 'wsg', 'LMA', 'CN', 'P', 'C'),
  trait = c('Leaf thickness', 'Leaf toughness', 'Leaf nitrogen mass', 'Wood specific gravity', 'Leaf mass per area', 'Leaf carbon per nitrogen', 'Leaf phosphorous', 'Leaf carbon'),
  TRYcode = c(46, NA, 57, 4, 11, 146, 58, 570)))
```

```{r species}
LES <- read.csv("~/Documents/ECOFOG/TROLL/inst/extdata/GLOPNET.csv", skip = 10)
LES <- LES[which(LES$BIOME == 'TROP_RF'),]
LES <- LES[which(LES$GF == 'T'),]
LES <- LES[-which(is.na(LES$log.LL)),]
LESsp <- unique(droplevels(LES$Species))
TRY <- read_delim("~/Documents/ECOFOG/TROLL/inst/extdata/TryAccSpecies.txt",
                  "\t", escape_double = FALSE, trim_ws = TRUE)
species <- data.frame(
  species = LESsp,
  TRYcode = TRY$AccSpeciesID[match(LESsp, TRY$AccSpeciesName)])
species <- species[-which(is.na(species$TRYcode)),]
rm(species, traits, LESsp, LES, TRY)
```
 
```{r data}
TRY <- read_delim("~/Documents/ECOFOG/TROLL/inst/extdata/TRY.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
TRY <- dcast(TRY, SpeciesName ~ TraitName, 
               value.var = 'StdValue', fun.aggregate = mean, na.rm = T)[-7]
names(TRY) <- c('Species', 'LMA', 'C', 'CN', 'Thick', 'wsg')
TRY$LMA <- 1/TRY$LMA*1000
data <- left_join(data, TRY, by = 'Species') ; rm(TRY)
data$LMA <- apply(data[c('LMA.x', 'LMA.y')], 1, mean, na.rm = T)
data <- data[-c(4,6)]
```

Table 2: Variable importance calculated with out-of the bag method applied on a random forest. *Leaf mass per area (LMA in $g.m^{-2}$), leaf nitrogen content (Nmass, in $mg.g^-1$) and leaf lifespan (LL in $months$) are taken in GLOPNET dataset from @wright_worldwide_2004. Leaf mass per area in TRY (LMAt in $g.m^{-2}$), leaf carbon content (C, in $mg.g^-1$) and leaf carbon per nitrogen (CN , in $g.g^-1$), leaf thickness (Thick in ) and wood specific gravity (wsg in ) are taken in TRY database (https://www.try-db.org).*

```{r OOB}
kable(randomForest(LL ~ LMA + Nmass + LMA + wsg, 
                   data = data, na.action = na.omit,
                   ntree = 1000, replace = F , importance = T,
                   do.trace = F, keep.forest = T, keep.inbag = T)$importance)
corrplot::corrplot.mixed(
  cor(data[c('LMA', 'Nmass', 'wsg')], 
      use = "pairwise.complete.obs"))
```

Figvure 2: Variable correlations. *Leaf mass per area (LMA in $g.m^{-2}$), leaf nitrogen content (Nmass, in $mg.g^-1$) and leaf lifespan (LL in $months$) are taken in GLOPNET dataset from @wright_worldwide_2004. Leaf mass per area in TRY (LMAt in $g.m^{-2}$), leaf carbon content (C, in $mg.g^-1$) and leaf carbon per nitrogen (CN , in $g.g^-1$), leaf thickness (Thick in ) and wood specific gravity (wsg in ) are taken in TRY database (https://www.try-db.org).*

If we try a new model **M10** with a $\beta_3$ to represent wood specific gravity *wsg*, we obtain following model:

$$ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j + {\beta_3}_s*{wsg_s}_j},\,\sigma)\,$$

```{r functions}
fitgraph <- function(fit){
  pars <- fit@model_pars
  beta <- pars[grep('beta', pars)]
  beta <- beta[-grep('s', beta)]
  sigma <- pars[grep('sigma', pars)]
  posterior <- as.matrix(fit)
  r <- mcmc_areas(posterior,  pars = c(beta, sigma, 'lp__'), prob = 0.8)
  c <- mcmc_trace(posterior,  pars = c(beta, sigma), n_warmup = 2000,
                   facet_args = list(nrow = 2, labeller = label_parsed))
  p <- ggpairs(data.frame(posterior[,beta]))
  return(list(r = r, c = c, p = p))
}
datasub <- function(data, dataset){
  LMAmax <- max(data$LMA)
  Nmax <- max(data$Nmass)
  wsgmax <- max(data$wsg)
  if(dataset != 'full')
    data <- data[-which(data$Dataset %in% dataset),]
  return(list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/LMAmax),
              Nmass = (data$Nmass/Nmax),
              wsg = (data$wsg/wsgmax)))
}
LLpred <- function(fit, data){
  LMA <- data$LMA
  N <- data$Nmass
  wsg <- data$wsg
  LMAmax <- max(data$LMA)
  Nmax <- max(data$Nmass)
  wsgmax <- max(data$wsg)
  pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),]
  beta <- pars[c('beta_0', 'beta_1', 'beta_2', 'beta_3')]
  beta[is.na(beta)] <- 0
  mu <- beta[1] + beta[2]*(LMA/LMAmax) - beta[3]*(N/Nmax) + beta[4]*(wsg/wsgmax)
  return(exp(mu))
}
evaluate <- function(fit, data){
  mean(sapply(levels(droplevels(data$Dataset)), function(dataset){
    fit1 <- fit[[dataset]]
    data1 <- data[which(data$Dataset == dataset),]
    fi <- LLpred(fit1, data1)
    yi <- data1$LL
    ei <- yi - fi
    rmse <- sqrt(mean(ei^2))
  }))
}
```

```{r model}
mpath <- "~/Documents/ECOFOG/TROLL/inst/doc/LL_models"
# rstan_options(auto_write = TRUE)
# options(mc.cores = detectCores())
# data <- data[c('Dataset', 'LL', 'LMA', 'Nmass', 'wsg')]
# data <- data[-which(is.na(data$wsg)),]
# data <- data[-which(is.na(data$LMA)),]
# LMAmax <- max(data$LMA) ; Nmax <- max(data$Nmass) ; wsgmax <- max(data$wsg)
# mdata <- sapply(c('full', levels(droplevels(data$Dataset))),
# function(x) datasub(data, x), simplify = F)
# fits <- lapply(as.list(paste0('M', c(1,10))), function(model){
#   fit <- lapply(mdata, function(x) stan(file = file.path(mpath, paste0(model, '.stan')), data = x))
#   eval <- evaluate(fit, data)
#   fit <- list(full = fit$full, eval = eval)
#   path <- file.path(mpath, paste0(model, '.LL.Rdata'))
#   save(fit, file = path)
# })
fits <- list()
i <- 1
models <- list.files(path = mpath, pattern = '.LL.Rdata')
for(file in models){
  load(file = file.path(mpath, file))
  fits[[i]] <- fit ; rm(fit) ; i <- i + 1
} 
names(fits) <- paste0('fit', c(10,1))
g <- fitgraph(fits$fit10$full) ; rm(file, i, models, mpath)
```

Maximum likekihood of **`r max(as.matrix(fits$fit10$full)[,'lp__'])`** for **M10** against **`r max(as.matrix(fits$fit1$full)[,'lp__'])`** for **M1** and RMSEP of **`r round(fits$fit10$eval,3)`** for **M10** against **`r round(fits$fit1$eval,3)`** for **M1**.

```{r r10}
g$r
```
Parameters posterior ditribution. *Light blue area represents 80% confidence interval, and vertical blue line the mean.*

```{r c10}
g$c
```
Parameters markov chains. *Light grey area represents warmup iterations.*

```{r p10}
g$p
```
Parameters pairs plot. *Parameters density distribution, pairs plot and Pearson's coefficient of correlation.*

# References
