---
title: "Leaf lifespan"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc())
library(flexdashboard)
library(knitr)
library(parallel)
library(TROLL)
library(RconTroll)
library(reshape2)
library(ggplot2)
library(cowplot)
library(entropart)
library(plotly)
library(rstan)
library(bayesplot)
library(rstanarm)
library(readr)
library(GGally)
cores <- detectCores()
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

Intro
==================

Issue {data-width=200}
-------------------------------------
    
TROLL model currently compute leaf lifespan with Reich's allometry [@Reich1991a]. But we have shown that Reich's allometry is underestimating leaf lifespan for low LMA species. Moreover simulations estimated unrealistically low aboveground biomass for low LMA species. We assumed Reich's allometry underestimation of leaf lifespan for low LMA species being the source of unrealistically low aboveground biomass inside TROLL simulations. We decided to find a better allometry with @wright_worldwide_2004 GLOPNET dataset.

We tested different models starting from complet model **Mcomp**:
$$ {LL_s}_j \sim \mathcal{logN}({\beta_0}*e^{{\beta_1}_s*{LMA_s}_j^{{\beta_3}_s} - {\beta_2}_s*{Nmass_s}_j^{{\beta_4}_s}},\,\sigma)\,$$

$$s=1,...,S_{=4}~, ~~j=1,...,n_s$$
$${\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}$$
We tested models M1 to M9 detailed in following tabs to find the better trade-off between:

1. Complexity (and number of parameters)
2. Convergence
3. Likelihood
4. Prediction quality with Root Mean Square Error (RMSE)

RMSE was computed by fitting a model without a dataset and evaluating it with the same dataset. RMSE in results are mean RMSE for each dataset. Results are shown for each models in eahc model tabs and summarized in Results tab.

LL graph {data-width=200}
-------------------------------------

```{r LL alo}
rm(list = ls())
LES <- read.csv("~/Documents/ECOFOG/TROLL/inst/extdata/GLOPNET.csv", skip = 10)
LES <- LES[which(LES$BIOME == 'TROP_RF'),]
LES <- LES[which(LES$GF == 'T'),]
LES <- LES[-which(is.na(LES$log.LL)),]
LES <- LES[-which(is.na(LES$log.LMA)),]
data <- cbind(LES[c('Dataset', 'Species')],
              10^LES[paste0('log.',c('LL', 'LMA', 'Nmass', 'Narea',
                                     'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea'))])
names(data) <- c('Dataset', 'Species', 'LL', 'LMA', 'Nmass', 'Narea',
              'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea')
data <- droplevels(data[c('Dataset', 'Species', 'LL', 'LMA', 'Nmass')])
g <- ggplot(droplevels(data), aes(x = LMA, y = LL, size = Nmass, color = Dataset, label = Species)) +
  geom_point(alpha = 0.5) + xlab('Leaf mass per area (LMA in g/m2)') +
  ylab('Leaf lifespan (LL in months)') + 
  scale_size_continuous(name = 'Leaf nitrogen content\n(Nmass, in mg/g)')
g ; rm(LES, g)
```

Figure 1: Leaf mass per area (LMA), leaf nitrogen content (Nmass) and leaflifespan (LL). *Leaf mass per area (LMA in $g.m^{-2}$), leaf nitrogen content (Nmass, in $mg.g^-1$) and leaf lifespan (LL in $months$) are taken in GLOPNET dataset from @wright_worldwide_2004.*

```{r functions}
fitgraph <- function(fit){
  pars <- fit@model_pars
  beta <- pars[grep('beta', pars)]
  beta <- beta[-grep('s', beta)]
  sigma <- pars[grep('sigma', pars)]
  posterior <- as.matrix(fit)
  r <- mcmc_areas(posterior,  pars = c(beta, sigma, 'lp__'), prob = 0.8)
  c <- mcmc_trace(posterior,  pars = c(beta, sigma), n_warmup = 2000,
                   facet_args = list(nrow = 2, labeller = label_parsed))
  p <- ggpairs(data.frame(posterior[,beta]))
  return(list(r = r, c = c, p = p))
}
datasub <- function(data, dataset){
  LMAmax <- max(data$LMA)
  Nmax <- max(data$Nmass)
  if(dataset != 'full')
    data <- data[-which(data$Dataset %in% dataset),]
  return(list(S = length(levels(droplevels(data$Dataset))), 
                          N = length(as.numeric(droplevels(data$Dataset))),
                          J = as.numeric(droplevels(data$Dataset)),
                          LL = data$LL,
                          LMA = (data$LMA/LMAmax),
                          Nmass = (data$Nmass/Nmax)))
}
LLpred <- function(fit, data){
  LMA <- data$LMA
  N <- data$Nmass
  LMAmax <- max(data$LMA)
  Nmax <- max(data$Nmass)
  pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),]
  beta <- pars[c('beta_0', 'beta_1', 'beta_2', 'beta_3', 'beta_4')]
  one <- (length(grep('Nmass', fit@stanmodel@model_code)) > 0)
  if(one)
    beta[is.na(beta)] <- 1
  if(!one)
    beta[is.na(beta)] <- 0
  beta[1] + beta[2]*(LMA/LMAmax)^beta[4] - beta[3]*(N/Nmax)^beta[5]
  return(exp(beta))
}
evaluate <- function(fit, data){
  mean(sapply(levels(droplevels(data$Dataset)), function(dataset){
    fit1 <- fit[[dataset]]
    data1 <- data[which(data$Dataset == dataset),]
    fi <- LLpred(fit1, data1)
    yi <- data1$LL
    ei <- yi - fi
    rmse <- sqrt(mean(ei^2))
  }))
}
```

```{r models, include=FALSE}
mpath <- "~/Documents/ECOFOG/TROLL/inst/doc/LL_models"
# rstan_options(auto_write = TRUE)
# options(mc.cores = detectCores())
# mdata <- sapply(c('full', levels(droplevels(data$Dataset))),
# function(x) datasub(data, x), simplify = F)
# fits <- lapply(as.list(paste0('M', 1:9, '.stan')), function(model){
#   fit <- lapply(mdata, function(x) stan(file = file.path(mpath, model), data = x))
#   eval <- evaluate(fit, data)
#   fit <- list(full = fit$full, eval = eval)
#   path <- file.path(mpath, paste0(model, '.Rdata'))
#   save(fit, file = path)
#   return(path)
# })
fits <- list()
i <- 1
for(file in list.files(path = mpath, pattern = '.stan.Rdata')){
  load(file = file.path(mpath, file))
  fits[[i]] <- fit ; rm(fit) ; i <- i + 1
} 
names(fits) <- paste0('fit', 1:9)
```

M1
==================

```{r m1, include=FALSE}
fit <- fits$fit1
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA - {\beta_2}_s*N,\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res1}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn1}
g$c ; g$p
```

M2
==================

```{r m2, include=FALSE}
fit <- fits$fit2
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + LMA^{{\beta_3}_s} - N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res2}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn2}
g$c ; g$p
```


M3
==================

```{r m3, include=FALSE}
fit <- fits$fit3
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA^{{\beta_3}_s} - {\beta_2}_s*N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res3}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn3}
g$c ; g$p
```


M4
==================

```{r m4, include=FALSE}
fit <- fits$fit4
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA - N,\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res4}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn4}
g$c ; g$p
```


M5
==================

```{r m5, include=FALSE}
fit <- fits$fit5
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + LMA^{{\beta_3}_s} - N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res5}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn5}
g$c ; g$p
```


M6
==================

```{r m6, include=FALSE}
fit <- fits$fit6
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA^{{\beta_3}_s} - N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res6}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn6}
g$c ; g$p
```


M7
==================

```{r m7, include=FALSE}
fit <- fits$fit7
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA,\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res7}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn7}
g$c ; g$p
```


M8
==================

```{r m8, include=FALSE}
fit <- fits$fit8
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + LMA^{{\beta_3}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res8}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn8}
g$c ; g$p
```


M9
==================

```{r m9, include=FALSE}
fit <- fits$fit9
g <- fitgraph(fit$full)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA^{{\beta_3}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit$full)[,'lp__'])`** and RMSE of **`r round(fit$eval,3)`**

```{r res9}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn9}
g$c ; g$p
```


Results
==================

Table {data-width=200}
-------------------------------------

Model **M1** seemed to have shown the best trade-off between complexity ($K=6$ parameters), convergence (see tab M1), likelihood, and prediction quality (see table 1). Figure 2 presents model prediction confidence interval and figure 3 compare Reich's allometry and model M1 predictions with species functional traits used in TROLL. **Nevertheless we will test other predictors gathered from TRY database following a model with the same response curve.**

Table 1: Models likelihood and prediction quality.*RMSE was computed by fitting a model without a dataset and evaluating it with the same dataset. RMSE in results are mean RMSE for each dataset.*

```{r table}
res <- data.frame(row.names = paste0('M', 1:9),
                  ML = unlist(lapply(fits, function(fit) round(max(as.matrix(fit$full)[,'lp__']),3))),
                  RMSE = unlist(lapply(fits, function(x) x$eval)))
kable(res)
fit <- fits$fit1$full ; rm(fits)
LMAmax <- max(data$LMA) ; Nmax <- max(data$Nmass)
beta <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('beta_0', 'beta_1', 'beta_2')]
```

$$LL = `r round(exp(beta[1]),3)`*e^{`r round(beta[2]/LMAmax,3)` *LMA -`r round(beta[3]/Nmax,3)`*Nmass }$$

Graphs {data-width=400}
-------------------------------------

```{r graph, fig.height=4}
LLpred <- function(LMA, Nmass, LMAmax, Nmax, beta0, beta1, beta2, sigma){
  mu <- beta0 + beta1*(LMA/LMAmax) - beta2*(Nmass/Nmax)
  rlnorm(1, mu, sigma)
}
pred <- apply(as.matrix(fit), 1, function(x)
  sapply(seq_len(dim(data)[1]), function(i)
      LLpred(data$LMA[i], data$Nmass[i], LMAmax, Nmax, x['beta_0'], x['beta_1'], x['beta_2'], x['sigma'])
    ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.05, 0.95)))))
names(pred) <- c('5%', '95%')
pred$mean <- exp((beta[1] + beta[2]*(data$LMA/LMAmax) - beta[3]*(data$Nmass/Nmax)))
pred$LMA <- data$LMA
pred$Nmass <- data$Nmass
g <- ggplot(data, aes(x = LMA, y = LL)) +
  geom_ribbon(aes(x = pred$LMA, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$LMA, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$LMA, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$LMA, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Leaf mass per area (LMA in g/m2)') +
  ylab('Leaf lifespan (LL in months)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
h <- ggplot(data, aes(x = Nmass, y = LL)) +
  geom_ribbon(aes(x = pred$Nmass, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$Nmass, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$Nmass, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$Nmass, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Leaf nitrogen content (Nmass in %)') +
  ylab('Leaf lifespan (LL in months)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
plot_grid(g, h, labels = c('LMA', 'Nmass'))
```

```{r TROLL, fig.height=4}
species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt', header=T, dec=".", sep="", row.names = 1)[c('LMA', 'Nmass')]
species$Reich <- 1.5 + 10^(7.18+3.03*log10(species$LMA*0.0001))
species$Model <- with(species, exp((beta[1] + beta[2]*(LMA/LMAmax) - beta[3]*(Nmass*100/Nmax))))
g <- ggplot(melt(species, id = c('LMA', 'Nmass')), aes(x = LMA, y = value, color = variable)) +
  geom_point() + theme_bw() +   xlab('Leaf mass per area (LMA in g/m2)') +
  ylab('Leaf lifespan (LL in months)')
h <- ggplot(melt(species, id = c('LMA', 'Nmass')), aes(x = Nmass, y = value, color = variable)) +
  geom_point() + theme_bw() +   xlab('Leaf nitrogen content (Nmass in %)') +
  ylab('Leaf lifespan (LL in months)')
plot_grid(g, h, labels = c('LMA', 'Nmass'))
```

References
==================

