---
title: "Leaf lifespan"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc())
library(flexdashboard)
library(knitr)
library(parallel)
library(TROLL)
library(RconTroll)
library(reshape2)
library(ggplot2)
library(cowplot)
library(entropart)
library(plotly)
library(rstan)
library(bayesplot)
library(rstanarm)
library(readr)
library(GGally)
cores <- detectCores()
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
    cache = T, cache.lazy = F)
```

Intro
==================

Issue {data-width=200}
-------------------------------------
    
TROLL model currently compute leaf lifespan with Reich's allometry [@Reich1991a]. But we have shown that Reich's allometry is underestimating leaf lifespan for low LMA species. Moreover simulations estimated unrealistically low aboveground biomass for low LMA species. We assumed Reich's allometry underestimation of leaf lifespan for low LMA species being the source of unrealistically low aboveground biomass inside TROLL simulations. We decided to find a better allometry with @wright_worldwide_2004 GLOPNET dataset.

We tested different models starting from complet model **Mcomp**:
$$ {LL_s}_j \sim \mathcal{logN}(log({\mu_s}_j),\,\sigma)\,, ~~s=1,...,S_{=4}~, ~~j=1,...,n_s$$

$${\mu_s}_j =  {\beta_0}*e^{{\beta_1}_s*{LMA_s}_j^{{\beta_3}_s} - {\beta_2}_s*{Nmass_s}_j^{{\beta_4}_s}}$$
$${\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I$$
$$(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}$$

LL graph {data-width=200}
-------------------------------------

```{r LL alo}
rm(list = ls())
LES <- read.csv("~/Documents/ECOFOG/TROLL/inst/extdata/GLOPNET.csv", skip = 10)
LES <- LES[which(LES$BIOME == 'TROP_RF'),]
LES <- LES[which(LES$GF == 'T'),]
LES <- LES[-which(is.na(LES$log.LL)),]
LES <- LES[-which(is.na(LES$log.LMA)),]
data <- cbind(LES[c('Dataset', 'Species')],
              10^LES[paste0('log.',c('LL', 'LMA', 'Nmass', 'Narea',
                                     'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea'))])
names(data) <- c('Dataset', 'Species', 'LL', 'LMA', 'Nmass', 'Narea',
              'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea')
data <- droplevels(data[c('Dataset', 'Species', 'LL', 'LMA', 'Nmass')])
g <- ggplot(droplevels(data), aes(x = LMA, y = LL, size = Nmass, color = Dataset, label = Species)) +
  geom_point(alpha = 0.5) + xlab('Leaf mass per area (LMA in g/m2)') +
  ylab('Leaf lifespan (LL in months)') + 
  scale_size_continuous(name = 'Leaf nitrogen content\n(Nmass, in mg/g)')
g ; rm(LES, g)
```

Figure 1: Leaf mass per area (LMA), leaf nitrogen content (Nmass) and leaflifespan (LL). *Leaf mass per area (LMA in $g.m^{-2}$), leaf nitrogen content (Nmass, in $mg.g^-1$) and leaf lifespan (LL in $months$) are taken in GLOPNET dataset from @wright_worldwide_2004.*

```{r functions}
fitgraph <- function(fit){
  pars <- fit@model_pars
  beta <- pars[grep('beta', pars)]
  beta <- beta[-grep('s', beta)]
  sigma <- pars[grep('sigma', pars)]
  posterior <- as.matrix(fit)
  r <- mcmc_areas(posterior,  pars = c(beta, sigma, 'lp__'), prob = 0.8)
  c <- mcmc_trace(posterior,  pars = c(beta, sigma), n_warmup = 2000,
                   facet_args = list(nrow = 2, labeller = label_parsed))
  p <- ggpairs(data.frame(posterior[,beta]))
  return(list(r = r, c = c, p = p))
}
```

```{r model, include=FALSE}
mpath <- "~/Documents/ECOFOG/TROLL/inst/doc/LL_models"
rstan_options(auto_write = TRUE)
options(mc.cores = detectCores())
```

M1
==================

```{r m1, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M1.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA + {\beta_2}_s*N,\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res1}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn1}
g$c ; g$p
```

M2
==================

```{r m2, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M2.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + LMA^{{\beta_3}_s} + N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res2}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn2}
g$c ; g$p
```


M3
==================

```{r m3, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M3.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA^{{\beta_3}_s} + {\beta_2}_s*N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res3}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn3}
g$c ; g$p
```


M4
==================

```{r m4, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M4.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA + N,\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res4}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn4}
g$c ; g$p
```


M5
==================

```{r m5, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M5.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + LMA^{{\beta_3}_s} + N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res5}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn5}
g$c ; g$p
```


M6
==================

```{r m6, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M6.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA^{{\beta_3}_s} + N^{{\beta_4}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res6}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn6}
g$c ; g$p
```


M7
==================

```{r m7, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M7.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA,\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res7}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn7}
g$c ; g$p
```


M8
==================

```{r m8, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M8.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + LMA^{{\beta_3}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res8}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn8}
g$c ; g$p
```


M9
==================

```{r m9, include=FALSE}
mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/max(data$LMA)),
              Nmass = (data$Nmass/max(data$Nmass)))
fit <- stan(file = file.path(mpath, 'M9.stan'), data = mdata)
g <- fitgraph(fit)
```

Model {data-width=300}
-------------------------------------

$$ LL \sim \mathcal{logN}(\beta_0 + {\beta_1}_s*LMA^{{\beta_3}_s},\sigma)\,$$
Maximum likekihood of **`r max(as.matrix(fit)[,'lp__'])`**

```{r res9}
g$r
```

Convergence {data-width=300} 
-------------------------------------

```{r diagn9}
g$c ; g$p
```


Result
==================

```{r final}
pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),]
LMA_max <- max(data$LMA)
Nmass_max <- max(data$Nmass)
beta_0 <- pars['beta_0']
beta_1 <- pars['beta_1']
beta_3 <- pars['beta_3']
LLpred <- function(LMA, LMA_max, beta_0, beta_1, beta_3){
  fact <- beta_0 + beta_1*(LMA/LMA_max)^beta_3
  unname(exp(fact))
}
LLpred_sigma <- function(LMA, LMA_max, beta_0, beta_1, beta_3, sigma){
  mu <- beta_0 + beta_1*(LMA/LMA_max)^beta_3
  rlnorm(1, mu, sigma)
}
```

Result {data-width=600}
-------------------------------------
$$LL = `r round(exp(beta_0),3)`*e^{`r round(beta_1/(LMA_max^beta_3),3)` *LMA^{`r round(beta_3,3)`}}$$

```{r graph}
pars <- as.matrix(fit)[-c(1:2000),]
pred <- apply(pars, 1, function(x) 
  sapply(data$LMA, function(y)
    LLpred_sigma(y, LMA_max, x['beta_0'], x['beta_1'], x['beta_3'], x['sigma'])
  ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.05, 0.95)))))
names(pred) <- c('5%', '95%')
pred$LMA <- data$LMA
g <- ggplot(data, aes(x = LMA, y = LL)) + 
  geom_ribbon(aes(x = pred$LMA, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$LMA, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$LMA, y = pred$`95%`, color = '95%'), linetype = 2) +
  stat_function(fun = function(x) LLpred(x, LMA_max, beta_0, beta_1, beta_3),
                aes(colour = "mean")) +
  geom_point() +
  theme_bw() +
  xlab('Leaf mass per area (LMA in g/m2)') +
  ylab('Leaf lifespan (LL in months)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
g
```

Figure 3: Model predictions.

References
==================

