---
title: "Sylviculture simulations"
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output:
  html_document: 
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()) ; invisible(gc())
library(knitr)
library(dplyr)
library(TROLL)
library(reshape2)
library(ggplot2)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 12,
    cache = T, cache.lazy = F)
path <- '~/Documents/ECOFOG/Results/disturbance/sylviculture/data/'
```

```{r harvest, echo=FALSE}
# harvest <- data.frame(
#   species = c('Brosimum_guianense', 'Brosimum_rubescens', 'Caryocar_glabrum',
#               'Dicorynia_guianensis', 'Goupia_glabra', 'Manilkara_bidentata',
#               'Manilkara_huberi', 'Ocotea_argyrophylla', 'Qualea_rosea', 
#               'Ruizterania_albiflora', 'Vouacapoua_americana',
#               'Couma_guianensis', 'Eperua_grandiflora', 'Lecythis_zabucajo',
#               'Moronobea_coccinea', 'Rhodostemonodaphne_grandis',
#               'Sterculia_pruriens', 'Sterculia_speciosa',
#               'Sterculia_villifera', 'Virola_michelii',
#               'Vochysia_guianensis', 'Bocoa_prouacensis',
#               'Couratari_multiflora', 'Eperua_falcata',
#               'Eperua_rubiginosa'),
#   dbh_min = 0.55,
#   dbh_max = 2,
#   interest = c(rep(1,11), rep(2,10), rep(3,4))
# )
# row.names(harvest) <- harvest$species
# harvest['Vouacapoua_americana','dbh_max'] <- 0.70
```

```{r mature, cache=TRUE, echo=FALSE, message=FALSE}
# mature <- loadStack('~/Documents/ECOFOG/Results/disturbance/maturity/models/')
# save(mature, file = '~/Documents/ECOFOG/Results/disturbance/maturity/mature.Rdata')
# load('~/Documents/ECOFOG/Results/disturbance/maturity/mature.Rdata')
# data <- data.frame(species = c('Total'))
# for(layer in mature@layers){
#   cat(layer@name, '\n')
#   data_l <- layer@final_pattern@data
#   data_l$species <- row.names(layer@sp_par)[data_l$sp_lab]
#   data_l <- data_l %>%
#     left_join(harvest, by = 'species') %>%
#     filter(!is.na(dbh_min)) %>%
#     filter(dbh >= dbh_min) %>%
#     filter(dbh <= dbh_max) %>%
#     mutate(volume = -0.0358 + 8.7634*dbh^2) %>%
#     group_by(species) %>%
#     summarise_each(funs(sum)) %>%
#     select(species, volume)
#   names(data_l)[2] <- layer@name
#   data <- full_join(data, data_l, by = 'species')
# }
# row.names(data) <- data$species
# data <- data[-1]
# data[is.na(data)] <- 0
# data['Total',] <- colSums(data)
# data <- t(data)
# save(data, file = '~/Documents/ECOFOG/Results/disturbance/maturity/matureSylv.Rdata')
# load('~/Documents/ECOFOG/Results/disturbance/maturity/matureSylv.Rdata')
# data <- data.frame(data)
# data <- data[-which(data$Total == 0),]
# names(data) <- gsub('_', '\n', names(data))
# data <- round(data/16, 3)
# DT::datatable(data, 
#               options = list(pageLength = 47, order = c(1, 'desc')),
#               caption = 'Harvestable volume (m3/ha)')
```

```{r d Sim, include=FALSE}
# rm(data, harvest)
# sim <- row.names(data)
# # Outputs
# path <- '~/Documents/ECOFOG/Results/disturbance/sylviculture'
# dir.create(file.path(path))
# invisible(sapply(sim, function(s) dir.create(file.path(path, s))))
# # Inits
# invisible(sapply(sim, function(s){
#   file.copy(from = file.path(
#     '~/Documents/ECOFOG/Results/disturbance/maturity/', 
#     'parameters', paste0(s, '.txt')),
#     to  = file.path(path, paste0('init_', s, '.txt')))}))
# # Forests
# load('~/Documents/ECOFOG/Results/disturbance/maturity/mature.Rdata')
# invisible(lapply(sim, function(s){
#   layer <- mature@layers[[s]]
#   cat(layer@name, '\n')
#   inventoryFromOutput(layer, paste0('forest_', layer@name, '.txt'), path)
# }))
# # Sylviculture
# invisible(lapply(sim, function(s){
#   layer <- mature@layers[[s]]
#   cat(layer@name, '\n')
#   sylvicultureInit(layer@sp_par, 
#                    paste0('sylviculture_', layer@name, '.txt'), 
#                    path)
# }))
# # Bash
# cores <- 20
# if('sylviculture.sh' %in% list.files(file.path(path)))
#    unlink(file.path(path, 'sylviculture.sh'))
# bash <- file(file.path(path, 'sylviculture.sh'), open = 'a')
# cat(' #!/bin/bash  \n\n', file = bash, append = TRUE)
# cat(' echo "sylviculture forest parrallel modelling in TROLL" \n\n', file = bash, append = TRUE)
# for(i in seq_len(length(sim))){
#   command <- paste0('./troll',
#                     ' -i"init_', sim[i], '.txt"',
#                     ' -f"forest_', sim[i], '.txt"',
#                     ' -s"sylviculture_', sim[i], '.txt"',
#                     ' -o./', sim[i], '/', sim[i], ' &\n')
#   cat(command, file = bash, append = TRUE)
#   if(i %/% cores == i/cores)
#   cat('wait\n', file = bash, append = TRUE)
# }
# cat('wait\n', file = bash, append = TRUE)
# cat(' echo "TROLL all jobs done"  \n\n', file = bash, append = TRUE)
# close(bash)
```

# Results

```{r tracks}
# # Load stack 
# sylviculture <- loadStack('~/Documents/ECOFOG/Results/disturbance/sylviculture/models/')
# save(sylviculture, file = file.path(path, 'sylviculture.Rdata'))
# load('~/Documents/ECOFOG/Results/disturbance/sylviculture/sylviculture.Rdata')
# tracks <- lapply(sylviculture@layers, function(layer)
#                  layer@disturbance %>% 
#     ggplot(aes(col, row, color = type, size = dbh)) + 
#     geom_point() + xlim(0,400) + ylim(0,400))
# save(doe, file = '~/Documents/ECOFOG/Results/disturbance/disturbance/disturbanceDOE.Rdata')
# save(tracks, file = file.path(path, 'tracks.Rdata'))
```

```{r DOE}
# ba <- data.frame(unlist(lapply(sylviculture@layers, function(layer)
#   sum(pi*layer@disturbance$dbh^2/4))))
# names(ba) <- 'ba'
# ba$sim <- row.names(ba)
# load(file.path('~/Documents/ECOFOG/Results/disturbance/disturbance/data/disturbanceDOE.Rdata'))
# doe <- doe %>% 
#   select(-sim, -disturbance) %>% 
#   mutate(sim = paste0(richness, '.', cvh))
# doe <- left_join(ba, doe) %>% 
#   distinct()
# rm(ba)
# save(doe, file = file.path(path, 'DOE.Rdata'))
load(file.path(path, 'DOE.Rdata'))
```

## Ecosystem functions

```{r ecosystem}
# load('~/Documents/ECOFOG/Results/disturbance/sylviculture/sylviculture.Rdata')
# ecosystemS <- extractData(sylviculture)
# load(file.path('~/Documents/ECOFOG/Results/disturbance/disturbance/data/disturbanceData.Rdata'))
# ecosystem <- ecosystem %>%
#   filter(sim %in% paste0(unique(ecosystemS$sim), '.', 0)) %>%
#   bind_rows(ecosystemS, .)
# rm(ecosystemS)
# save(ecosystem, file = file.path(path, 'ecosystem.Rdata'))
load(file.path(path, 'ecosystem.Rdata'))
```

```{r resilience}
# sim <- unique(doe$sim)
# time <- unique(ecosystem$time)
# resilience <- sapply(sim, function(s){
#   data_s <- ecosystem[which(ecosystem$sim %in% c(s, paste0(s, '.0'))),]
#   data_s <- sapply(time, function(t){
#     data_t <- subset(data_s, time == t)
#     T0 <- which(data_t$sim == paste0(s, '.0'))
#     data_t[-c(1:2)] <- apply(data_t[-c(1:2)], 2, function(x) x/x[T0])
#     return(data_t[-T0,])
#   }, simplify = F)
#   data_s <- do.call('rbind', data_s)
# }, simplify = F) ; rm(sim, time)
# resilience <- do.call('rbind', resilience)
# save(resilience, file = file.path(path, 'resilience.Rdata'))
load(file.path(path, 'resilience.Rdata'))
```

```{r Deq}
deq <- data.frame(
  time = resilience$time,
  sim = resilience$sim,
  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2)))
)
```

```{r Ieq}
sim <- unique(resilience$sim)
Ieq <- sapply(sim, function(s){
  data_s <- subset(deq, sim ==s)
  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
}, simplify = F) ; rm(sim)
Ieq <- data.frame(do.call('rbind', Ieq))
names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
```

```{r gIeq, echo=FALSE, fig.cap='Figure 1: Ecosystem resilience after 600 years with taxonomic and functional diversity. Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years normalized by disturbed basal area was represented against functional richness [FRic, @villeger_new_2008]. Dot color represents the species richness (nb) whereas dot size represents the disturbed basal area (m2/ha). Grey line represents the linear regression and grey area the confidence interval.', cache=TRUE}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>% 
  ggplot(aes(x = FRic, y = Ieq.structure/ba, label = sim)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(color = richness, size = ba/16)) +
  xlab('Functional richness (FRic)') +
  ylab('Cummulative integral from ecosystem 
       distance to forest dynamic equilibrium after 600 years') +
  scale_color_discrete('Species\nrichness\n(SR)') +
  scale_size_continuous('Basla\narea\ndisturbed\n(m2/ha)')
```

```{r lmIeq, results='asis'}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>%
  select(Ieq.structure, ba, FRic, FEve, FDiv, FDis) %>% 
  lm(Ieq.structure/ba ~ FRic + FEve + FDiv + FDis, data = .) %>% 
  summary() %>% 
  pander::pander()
```

## Biodiversity effect

```{r NE}
# load('~/Documents/ECOFOG/Results/disturbance/maturity/mature.Rdata')
# load('~/Documents/ECOFOG/Results/disturbance/sylviculture/sylviculture.Rdata')
# load('~/Documents/ECOFOG/Results/disturbance/monoculture/disturbance/md.Rdata')
# BE <- lapply(sylviculture@layers, function(layer)
#   {cat(layer@name, '\n') ; return(getBE(layer, mature, md))})
# BE <- do.call('rbind', BE)
# BEs <- BE
# load('~/Documents/ECOFOG/Results/disturbance/disturbance/data/disturbanceBE.Rdata')
# BE <- BE %>%
#   filter(sim %in% paste0(unique(BEs$sim), '.', 0)) %>%
#   bind_rows(BEs, .)
# rm(BEs)
# save(BE, file = file.path(path, 'BE.Rdata'))
load(file.path(path, 'BE.Rdata'))
```

```{r BEresilience}
# sim <- as.character(unique(Ieq$sim))
# time <- unique(BE$time)
# be <- c('NE', 'CE', 'SE')
# BEresilience <- sapply(sim, function(s){
#   cat(s, '\n')
#   data_s <- BE[which(BE$sim %in% c(s, paste0(s, '.0'))),]
#   data_s <- sapply(time, function(t){
#     data_t <- subset(data_s, time == t)
#     T0 <- which(data_t$sim == paste0(s, '.0'))
#     T0_NE <- which(data_t[T0,]$BE == "NE")
#     data_t[-c(1:3)] <- apply(data_t[-c(1:3)], 2, function(x) x/x[T0_NE])
#     return(data_t[-T0,])
#   }, simplify = F)
#   data_s <- do.call('rbind', data_s)
# }, simplify = F) ; rm(sim, time, BE)
# BEresilience <- do.call('rbind', BEresilience)
# save(BEresilience, file = file.path(path, 'BErecovery.Rdata'))
load(file.path(path, 'BErecovery.Rdata'))
```

```{r gBE, echo=FALSE, fig.cap='Figure 4: Resilience of complementarity and selection effects. Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time.', cache=TRUE}
BEresilience %>%
  left_join(., doe, by = 'sim') %>%
  filter(BE %in% c('SE', 'CE')) %>%
  group_by(time, BE) %>%
  summarise(qt1 =quantile(agb, probs=0.05),
            qt2 =quantile(agb, probs=0.75),
            mean=mean(agb),
            n=n()) %>%
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  xlab('Time (years)') +
  ylab('Recovery of aboveground biomass net effect') +
  ylim(-2.5,2.5)
```

# Discussion

# Supplementary materials

## S1: Ecosystem functions

Supplementary materials S1 presents ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. It encompass all functionnal diversity components [FRIC, FEve, FDiv, and FDis, @villeger_new_2008]. And it presents results for both forest dynamic (Figure S1.1) and forest production (Figure S1.2).

```{r Ieq.structure all var, echo=FALSE, fig.cap='Figure S1.1: Ecosystem resilience after 600 years with taxonomic and functional diversity. Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years normalized by disturbed basal area was represented against functional functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008]. Dot color represents the species richness (nb) whereas dot size represents the disturbed basal area (m2/ha). Grey line represents the linear regression and grey area the confidence interval.', cache=TRUE}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>%
  select(Ieq.structure, ba, richness, FRic, FEve, FDiv, FDis) %>% 
  melt(id.vars = c('Ieq.structure', 'ba', 'richness')) %>% 
  ggplot(aes(x = Ieq.structure/ba, y = value)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(color = richness, size = ba/16)) +
  facet_wrap(~variable, scales = "free", labeller = 'label_both') + 
  ylab('Functional diversity') +
  xlab('Cummulative integral from ecosystem 
       distance to forest dynamic equilibrium after 600 years') +
  scale_color_discrete('Species\nrichness\n(SR)') +
  scale_size_continuous('Basla\narea\ndisturbed\n(m2/ha)')
```

```{r Ieq.production all var, echo=FALSE, fig.cap='Figure S1.2: Ecosystem resilience after 600 years with taxonomic and functional diversity. Cummulative integral from ecosystem distance to forest functionning equilibrium after 600 years normalized by disturbed basal area was represented against functional functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008]. Dot color represents the species richness (nb) whereas dot size represents the disturbed basal area (m2/ha). Grey line represents the linear regression and grey area the confidence interval.', cache=TRUE}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>%
  select(Ieq.production, ba, richness, FRic, FEve, FDiv, FDis) %>% 
  melt(id.vars = c('Ieq.production', 'ba', 'richness')) %>% 
  ggplot(aes(x = Ieq.production/ba, y = value)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(color = richness, size = ba/16)) +
  facet_wrap(~variable, scales = "free", labeller = 'label_both') + 
  ylab('Functional diversity') +
  xlab('Cummulative integral from ecosystem 
       distance to forest dynamic equilibrium after 600 years') +
  scale_color_discrete('Species\nrichness\n(SR)') +
  scale_size_continuous('Basla\narea\ndisturbed\n(m2/ha)')
```

## S2: Biodiversity effect

Supplementary materials S2 presents the resilience of complementarity and selection effects for different ecosystem metrics (AGB, BA, N, GPP and NPP, see Figure S2.1)

```{r allBE, echo=FALSE, fig.cap='Figure S2.1: Resilience of complementarity and selection effects. Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time for different ecosystem variables (AGB, BA, N, GPP).', cache=TRUE}
BEresilience %>% 
  select(BE, time, sim, agb, ba, n, gpp) %>% 
  melt(id.vars = c('BE', 'time', 'sim')) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(BE %in% c('SE', 'CE')) %>% 
  group_by(time, BE, variable) %>%
  summarise(qt1 =quantile(value, probs=0.05, na.rm = T),
            qt2 =quantile(value, probs=0.75, na.rm = T),
            mean=mean(value, na.rm = T)) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_wrap(~ variable, scales = "free", labeller = 'label_both') +
  xlab('Time (years)') +
  ylab('Recovery of variable net effect') + 
  ylim(-2.5,2.5)
```

# References
